Supporting Information

### 2018-03-30 SupportingInformation.Rmd
various changes

## Aphid data \label{Aphid_data}
   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
   * We threw out all locations that had fewer than 30 years data 
   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data

## North Sea plankton data \label{Pns_data}
   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
   * We threw out all locations that had fewer than 45 years data
   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data

@@

## Results: aphid abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

@@

## Results: aphid first flight data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}

@@

<!--Dan stopped editing here 2018 03 30-->

## North sea plankton data for 'Ceratium furca' \label{Ex_pns}

@@

```{r read_plankton_north_sea_data, echo=F}
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
```

```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"), mtime("FittingCopula_selective_loc.R"))}

@@

@@

\newpage
```{r tab_pns_AICw,echo=F,results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
             format="pandoc",
             digits=3,
             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
             booktabs=TRUE)
```

\newpage
```{r tab_pns_p_CvM,echo=F,results="markup"}
knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
             format="pandoc",
             digits=3,
             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
             booktabs=TRUE)
```

@@

Table \ref{tab_pns_bestcop} shows that several copulas are much better supported than the Gaussian copula, answering the first part of question 1 for these data. Table \ref{tab_pns_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_pns_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_pns_p_KS}] of the goodness of fit with atleast
`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_pns_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_pns_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_pns_LTdep_AICw} and Table \ref{tab_pns_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_pns_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`** location pair more lower tail dependence.

>>>>

## Abundance data for the green spruce aphid and first flight date data for the leaf-curling plum aphid in the UK\label{Aphid_data}
   * Aphid data contains 3 measurements [count, firstflight, flight-duration] at 11 locations for 20 species and for 35 years.
   * We threw out all locations that had fewer than 30 years data 
   * We use species=10 [Name : "Green spruce aphid"] for aphid-count data 
   * We use species=11 [Name : "Leaf-curling plum aphid"] for aphid-firstflight data

## Abundance data for *Ceratium furca* in UK seas \label{Pns_data}
   * Plankton North sea dataset collects data at 26 locations for 22 species and for 56 years.
   * We threw out all locations that had fewer than 45 years data
   * We use species=16 [Name : "Ceratium_furca"] for Plankton_North_Sea data
   * *Ceratium furca* is a dicoflagellate commonly found in UK seas
   * note, when we edit this, we cannot call it the North Seas since data come from around the UK not only the North Sea, unless Shya threw out all grid cells that were outside the North Sea

@@

## Results: abundance data for the green spruce aphid, *Elatobium abietinum* \label{Ex_count}

@@

## Results: first flight date data for the leaf-curling plum aphid, *Brachycaudus helichrysi* \label{Ex_ff}

@@

## Results: abundance data for *Ceratium furca* in UK seas \label{Ex_pns}

@@

<!--Load the data-->
```{r read_plankton_north_sea_data, echo=F}
d0_plankton_north_sea<-as.matrix(read.csv("./Data/Plankton_North_Sea_data/planktontimeseries201117_4.csv",header=F))
data_plankton_north_sea<-d_allsp_data_plankton_north_sea(d0_plankton_north_sea)
```

<!--Do the main analysis-->
```{r RES_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("OurBiCopSelect.R"),mtime("MyBiCopGofTest.R"), mtime("FittingCopula_selective_loc.R"))}

@@

As in previous sections (see also Methods) model selection among copulas was done separately for 
each pair of sampling locations. There were $14$ sampling locations, so results are presented in 
$14 \times 14$ matrix format, with rows and columns corresponding to locations. 
Table \ref{tab_pns_bestcop} shows that, for the large majority of location pairs, some 
non-normal copula was a better fit, according to AIC, than was the normal copula, answering the 
first part of question 1 for these data. Comparing the AIC weight of the lowest-AIC copula 
family for each pair of sampling locations (Table \ref{tab_pns_AICw}) to the AIC weight of
the normal copula family for the same pair (Table \ref{tab_pns_AICw_normal})

 Table \ref{tab_pns_AICw} represents the Akaike weights of the best-fit copula for each location pair. Our fitting results are supported by the lowest-AIC copula *p*-value in the Cramer-von Mises-based test [Table \ref{tab_pns_p_CvM}] of the goodness of its fit, and in the Kolmogorov-Smirnov-based test [Table \ref{tab_pns_p_KS}] of the goodness of fit with atleast
`r min((RES_plankton_north_sea$gfc_numBS_success/RES_plankton_north_sea$gfc_numBS),na.rm=T)*100` % successful bootstrap. Table \ref{tab_pns_AICw_normal} represents the Akaike weights of the normal copula for each location pair and by comparing with Table \ref{tab_pns_AICw}, one can verify that for most of the locations a normal copula is not a good fit based on minimum AIC. The model-averaged (using AIC) lower-tail dependence statistic and the model-averaged (again, AIC based) upper-tail dependence statistic are presented in Table \ref{tab_pns_LTdep_AICw} and Table \ref{tab_pns_UTdep_AICw} respectively, and these values are distinct from what a Gaussian copula would give (i.e., 0), answering the second part of question 1 for these data. Furthermore, the difference of these two quantities (lower minus upper) in Table \ref{tab_pns_LTmUT} indicates it is different from $0$ for most of the locations, answering the third part of question 1. These results do convincingly show non-Gaussian copula structure, having **`r sum((RES_plankton_north_sea$LTdep_AICw-RES_plankton_north_sea$UTdep_AICw)>0,na.rm=T)`** out of **`r (dim(RES_plankton_north_sea$gfc_p_CvM)[1]^2-dim(RES_plankton_north_sea$gfc_p_CvM)[1])-2*RES_plankton_north_sea$num_indep_loc_pair`** location pair more lower tail dependence.

@@

```{r tab_pns_AICw,echo=F,results="markup"}
knitr::kable(RES_plankton_north_sea$info_ord_AICw[,,1],
             format="pandoc",
             digits=3,
             caption = "Plankton_north_sea : AICw for best fit copula family for each selective location pair\\label{tab_pns_AICw}",
             booktabs=TRUE)
```

```{r tab_pns_p_CvM,echo=F,results="markup"}
knitr::kable(RES_plankton_north_sea$gfc_p_CvM,
             format="pandoc",
             digits=3,
             caption = "Plankton_north_sea : p-values (Cramer-von Mises-based test) for lowest AIC based best fit copula\\label{tab_pns_p_CvM}",
             booktabs=TRUE)
```

@@
